---
title: "salmonIPM: An integrated population model for Pacific salmon"
output:
  html_document:
    fig_caption: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r set_options, echo=FALSE, cache=FALSE}
options(width = 100)
```

***

### Contributors

__Eric R. Buhle__  
_Ocean Associates, Northwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration, Seattle, WA USA, eric.buhle@noaa.gov_

[__Mark D. Scheuerell__](https://faculty.washington.edu/scheuerl/)  
_Fish Ecology Division, Northwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration, Seattle, WA USA, mark.scheuerell@noaa.gov_

[__James T. Thorson__](https://sites.google.com/site/thorsonresearch/)  
_Fishery Resource Analysis and Monitoring Division, Northwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration, Seattle, WA USA, james.thorson@noaa.gov_

### Version

This is version `r paste0('0.',format(Sys.time(), '%y.%m.%d'))`.

***

## Background

`salmonIPM` is a modification of the [ASSESSOR](https://github.com/mdscheuerell/ASSESSOR) model, which was developed for only one population of Pacific salmon. New developments include:

1. the optional ability to model multiple populations simultaneously in a hierarchical framework;
2. allowing hatchery-origin adults to contribute to the spawning population, and distinguishing them from natural-origin adults; 
3. including removals of natural-origin adults due to hatchery broodstock collection in addition to harvest; 
4. modeling age composition as additive log-ratios rather than as a Dirichlet distribution; and
5. use of Hamiltonian Monte Carlo, as implemented in the [Stan](http://mc-stan.org/) programming language, to estimate all parameters and states.

## Requirements

At a minimum, `salmonIPM` requires the following data:

1. observed total number of adult spawners (escapement) by population and year;  
2. observed age composition of adult spawners by population and year (frequencies by age, based on finite samples);  
3. observed proportion of total spawners that were of hatchery origin (frequencies of each rearing origin, based on finite samples);  
4. observed numbers of adults removed for broodstock by population and year; and  
5. observed total harvest mortality by population and year.

Optionally, time series of environmental conditions may also be used as covariates of annual cohort productivity. In the multi-population case, these covariates are assumed to influence all populations identically. 

## Model structure

`salmonIPM` fits a hierarchical model with separate process and observation components.

### Process component

The basis for the process component is a function describing the density-dependent production and survival of offspring from their parents.

### Observation component


## User inputs

```{r get_user_inputs}
## URL for example data files
## set to NULL if using a local folder/directory
ex_url <- "https://raw.githubusercontent.com/ebuhle/master/IPM_PVA/"
```

